---
title: "Stat 311 Final Project"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2022-12-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("r package", repos = "http://cran.us.r-project.org")

```

```{r}
install.packages("tidyverse")
install.packages("scales")
library(tidyverse)
```



```{r message=FALSE, warning=FALSE, include=FALSE}
#install.packages("leaps")
remove.packages("rlang")
install.packages("scales") 
install.packages("rlang")
#install.packages("LibPath")
install.packages("DAAG")
install.packages("tidyverse")
#install.packages("ggplot2")
#install.packages("randomForest")
#install.packages("ISLR2")
#install.packages("caret")
#install.packages("gbm")
library(leaps)
library(lubridate)
library(MASS)
library(ISLR2)
library(DAAG)
library(dplyr)
library(tidyverse)
library(randomForest)
library(gbm)
```



```{r}
#Reading the dataset and changing the volume to numeric.
Tesla <- read.csv("C:/Users/Kunal/Desktop/STAT 311 Final Project/TSLA.csv")

Tesla$Volume = as.numeric(Tesla$Volume)/1e6
Tesla$Date <- as.Date(Tesla$Date)
str(Tesla) 
```



```{r}
png(file="./Tesla12.png");Tesla %>% ggplot( aes(x=Date, y=Adj.Close)) +
    geom_line() +
    geom_point(); dev.off()

summary(Tesla)
```



```{r}

png(file="./TeslaVol.png");Tesla %>% filter(Date>"2009-01-01") %>% group_by(month = lubridate::floor_date(Date, 'month')) %>% group_by(month) %>% summarise(mean_vol= mean(Volume)) %>% filter(month>"2020-01-01") %>% ggplot() +geom_bar( aes(x=month, y=mean_vol),stat="identity",colour="#006000") + labs( x= "Months", y="Monthly Average" ,title = "Monthly Average Traded Volume since Covid (in Millions)"); dev.off()


```





```{r}
Tesla %>% filter(Date>"2022-06-01") %>% ggplot() +geom_line( aes(x=Date, y=High),stat="identity",colour="#006000") + labs( x= "Date", y="Daily High Stock Price" ,title = "Daily Highs for last 6 months") 
```



```{r}

colnames(Tesla)
Cols =  colnames(Tesla)[colnames(Tesla)!="Date"]
Cols

round(cor(Tesla[Cols]), digits=3)

```



```{r}
fit = lm(High~Adj.Close, data = Tesla)
png(file="./TeslaHigh.png");plot(High~Adj.Close, data=Tesla, col="dodgerblue", main ="Daily High versus Adjusted Closing Price since 2010",xlab="Adjusted Closign Price",ylab="HIghest Stock Price of that Day")
abline(fit, col="green", lty=2, lwd=3); dev.off()
```



```{r}
print("value of R^2, adjusted R^2, AIC and BIC for the fitted model:")
c( summary(fit)$r.squared, 
  summary(fit)$adj.r.squared,
  AIC(fit),
  BIC(fit) )

```


# Model Selection for Predicting


```{r}
# Model 1 using Shots
mod1 = lm(log(High)~Adj.Close, data=Tesla)
plot(mod1$res~mod1$fitted)
#get summary and R-Squared
summary(mod1)
c(summary(mod1)$r.squared, summary(mod1)$adj.r.squared)
```

```{r}
# Model 2 using Open price
mod2 = lm(log(High)~Open, data=Tesla)
plot(mod2$res~mod2$fitted)
#get summary and R-Squared
summary(mod2)
c(summary(mod2)$r.squared, summary(mod2)$adj.r.squared)
```




```{r}
# Model 2 using High price
mod3 = lm(log(High)~Volume, data=Tesla)
plot(mod3$res~mod3$fitted)
#get summary and R-Squared
summary(mod3)
c(summary(mod3)$r.squared, summary(mod3)$adj.r.squared)
```



```{r}
# Model 2 using Open price
mod4 = lm(log(High)~Low, data=Tesla)
plot(mod4$res~mod4$fitted)
#get summary and R-Squared
summary(mod4)
c(summary(mod4)$r.squared, summary(mod4)$adj.r.squared)
```





```{r}
# Model 2 using Open price
mod5 = lm(log(High)~Adj.Close+Open+Volume+Low, data=Tesla)
plot(mod5$res~mod5$fitted)
abline(0,0,col = "red")
#get summary and R-Squared
summary(mod5)
c(summary(mod5)$r.squared, summary(mod5)$adj.r.squared)
```

```{r}
# model with all subset
mod.allsubset = regsubsets(log(High)~Adj.Close+Open+Volume+Low, data=Tesla)
summary(mod.allsubset)
```

```{r}
summary(mod.allsubset)$rsq # R-squared
summary(mod.allsubset)$adjr2 # adjusted R-squared
summary(mod.allsubset)$cp # Cp
summary(mod.allsubset)$bic #BIC
```

```{r}
par(mfrow = c(2, 2))
plot(summary(mod.allsubset)$rsq, xlab = "Number of Variables",
    ylab = "R^2", type = "l")
plot(summary(mod.allsubset)$adjr2, xlab = "Number of Variables",
    ylab = "Adjusted R^2", type = "l")
plot(summary(mod.allsubset)$cp, xlab = "Number of Variables",
    ylab = "Cp", type = "l")
plot(summary(mod.allsubset)$bic, xlab = "Number of Variables",
    ylab = "BIC", type = "l")
# Tip: type = "l" connects lines between the plotted 6 points.
```


```{r}
r_max = which.max(summary(mod.allsubset)$adjr2)
plot(summary(mod.allsubset)$adjr2, xlab = "Number of Variables",
    ylab = "Adjusted RSq", type = "l")
points(r_max, summary(mod.allsubset)$adjr2[r_max], col = "red", cex = 2, 
    pch = 20)
```



```{r}
bic_min = which.min(summary(mod.allsubset)$bic)
plot(summary(mod.allsubset)$bic, xlab = "Number of Variables",
    ylab = "BIC", type = "l")
points(bic_min, summary(mod.allsubset)$bic[bic_min], col = "red", cex = 2, 
    pch = 20)
```




```{r}
fit1 = lm(log(High)~Adj.Close+Open+Volume+Low, data=Tesla)
fit2 = lm(log(High)~Adj.Close+Volume+Low, data=Tesla)
par(mfrow = c(1, 2))
png(file="./Teslafit1.png");plot(fit1$res~fit1$fitted)
abline(0,0,col = "red");dev.off()
plot(fit2$res~fit2$fitted)
abline(0,0,col = "red")
```



```{r}
eval_mat = matrix(0, nrow=3, ncol=4) # initiate a 3 by 3 matrix.
colnames(eval_mat) = c("R-sq","adj R-sq", "AIC", "BIC")
eval_mat[1,] = c( summary(fit)$r.squared, summary(fit)$adj.r.squared,AIC(fit),BIC(fit) )
eval_mat[2,] = c( summary(fit1)$r.squared, summary(fit1)$adj.r.squared,AIC(fit1),BIC(fit1) )
eval_mat[3,] = c( summary(fit2)$r.squared, summary(fit2)$adj.r.squared,AIC(fit2),BIC(fit2) )
print(eval_mat)
```
# Conclusion
#### fit is the best model based on 𝑅^2 , adjusted 𝑅^2.
#### fit1 is the best model based on AIC and BIC.



```{r}
ms.vec = NULL 
# initate a vector to record overall Mean Square Error (MSE) of each model
```

```{r}
cv.fit = CVlm(Tesla,fit, m=3)
```
```{r}
# Obtain the overall MSE value and record it
attr(cv.fit,"ms")
ms.vec = c(ms.vec, attr(cv.fit,"ms"))
```

```{r}
cv.fit1 = CVlm(Tesla,fit1, m=3)
attr(cv.fit1,"ms")
ms.vec= c(ms.vec, attr(cv.fit1,"ms"))
```


```{r}
cv.fit2 = CVlm(Tesla,fit2, m=3)
```

```{r}
attr(cv.fit2,"ms")
ms.vec= c(ms.vec, attr(cv.fit2,"ms"))
```

```{r}
ms.vec
```
# Conslusion
#### Based on parsimony or KISS rule, I personally would recommend fit1 which is the best model selected by the AIC, BIC and now the lowest MSE of 0.5378050 based on 3-fold CV.


```{r}
summary(fit1)
```
# TREE METHOD

```{r}
##Removing the date column from the dataset
Tesla_p <- Tesla[c(2:7)]
Tesla_p
```


```{r}
# 1 Data Partition and RMSE function
set.seed(310)
index = sample(1:nrow(Tesla_p), nrow(Tesla_p)*0.8, replace=F)
train_t = Tesla_p[index,]
test_t = Tesla_p[-index,]
dim(train_t)
dim(test_t)
```

```{r}
myRMSE = function(actual, predicts){
mse = mean( (actual-predicts)^2)
return( sqrt(mse) )
}
```

```{r warning=FALSE}
#1.2 Linear regression
mod.lin = lm(High~.,data=Tesla_p)
pred.train=predict(mod.lin, newdata=train_t)
pred.test = predict(mod.lin, newdata=test_t)
#find MSE on training and testing
lin.rmse.trt = myRMSE(train_t$High, pred.train)
lin.rmse.tst = myRMSE(test_t$High, pred.test)
cat("training RMSE:", lin.rmse.trt)
cat("\ntesting RMSE:", lin.rmse.tst)

```

```{r}
library(tree)
```

```{r}
atree = tree(High~., data = train_t)
summary(atree)
```

```{r}
plot(atree)
text(atree, pretty = 0)
title(main = "Unpruned Regression Tree")
```
# Coclusion
#### Variable Close is the closing price of the stock of that day while the variable open is opening price of the stock.
#### tree indicates larger value of closing or lower values of opening price, corresponds to a higher range of High Stock Price of that day. 


```{r}
# Do cross-validation to select a good pruning of the tree
set.seed(310)
tree_cv = cv.tree(atree, K=10)
plot(tree_cv$size, sqrt(tree_cv$dev/nrow(train_t)),
type="b",
xlab= "Tree Size",
ylab = "CV-RMSE")
```



```{r}
tree_prune = prune.tree(atree , best = 5)
summary(tree_prune)
#plot tree
plot(tree_prune)
text(tree_prune, pretty = 0)
title(main = "Pruned Regression Tree")
```

```{r}
tree.pred.train = predict(tree_prune, newdata = train_t)
tree.pred.test = predict(tree_prune, newdata = test_t)
# find MSE on training and testing
tree.rmse.trt = myRMSE(train_t$High,tree.pred.train)
tree.rmse.tst = myRMSE(test_t$High, tree.pred.test)
cat("training RMSE:", tree.rmse.trt)
cat("\ntesting RMSE:", tree.rmse.tst)
```

```{r}
plot(tree.pred.test, test_t$High,
xlab = "Predicted",
ylab = "Actual",
main = "Predicted vs Actual: Single Tree, Test Data",
col = "Blue", pch = 20)
grid()
abline(0,1, col="red", lty = 2, lwd = 2)
```
# Ensemble Tree Methodss



```{r}
# Bagging
mod.bag = randomForest(High ~ .,
                      data = train_t,
                      mtry = 5,
                      type= "regression",
                      importance = TRUE,
                      ntrees = 500)
mod.bag
```


```{r}
bag.pred.train = predict(mod.bag, newdata=train_t)
bag.pred.test = predict(mod.bag, newdata=test_t)
bag.rmse.trt = myRMSE(train_t$High, bag.pred.train)
bag.rmse.tst = myRMSE(test_t$High, bag.pred.test)
cat("training RMSE:", bag.rmse.trt)
cat("\ntesting RMSE:", bag.rmse.tst)
```

```{r}
plot(bag.pred.test,test_t$High,
xlab = "Predicted", ylab = "Actual",
main = "Predicted vs Actual: Bagged Model, Test Data",
col = "dodgerblue", pch = 20)
grid()
abline(0,1, col ="red", lty = 2, lwd = 2)
```

```{r}
forest = randomForest(High~.,
                      data = train_t,
                      mtry = 2,
                      type="regression",
                      importance = TRUE,
                      ntrees = 500)
forest
```

```{r}
forest.pred.train = predict(forest,newdata=train_t)
forest.pred.test = predict(forest,newdata=test_t)
forest.rmse.trt = myRMSE(train_t$High,forest.pred.train)
forest.rmse.tst = myRMSE(test_t$High, forest.pred.test)
cat("training RMSE:", forest.rmse.trt)
cat("\ntesting RMSE:", forest.rmse.tst)
```

```{r}
#OOB rate
myRMSE(forest$predicted,train_t$High)
```

```{r}
png(file="./TeslaForest.png");plot(forest.pred.test, test_t$High,
xlab= "Predicted",
ylab = "Actual",
main = "Predicted vs Actual: Random Forest, Test Data",
col = "dodgerblue",pch=20)
grid()
abline(0,1,col="red", lty = 2, lwd = 2); dev.off()
```

```{r}
#type: 1 = mean decrease in accuracy
# 2 = mean decrease in node impurity
importance(forest, type = 1)
```

```{r}
varImpPlot(forest, type = 1)
```

```{r}
boost = gbm(High~.,
            data = train_t,
            distribution = "gaussian",
            n.trees = 5000,
            interaction.depth = 4,
            shrinkage= 0.1)
boost
```

```{r}
boost.pred.train = predict(boost,newdata=train_t,n.trees = 5000)
boost.pred.test = predict(boost,newdata=test_t,n.trees = 5000)
boost.rmse.trt = myRMSE(train_t$High,boost.pred.train)
boost.rmse.tst = myRMSE(test_t$High, boost.pred.test)
cat("training RMSE:", boost.rmse.trt)
cat("\ntesting RMSE:", boost.rmse.tst)
```


```{r}
png(file="./Teslasummary.png");summary(boost);dev.off()
```

```{r}
plot(boost, i = "Open", col = "dodgerblue", lwd = 2)
plot(boost, i = "Low", col = "dodgerblue", lwd = 2)
png(file="./TeslaClose.png");plot(boost, i = "Close", col = "dodgerblue", lwd = 2);dev.off()
```

```{r}
as_tibble(cbind( method=c("Lin Regression","Pruned Tree","Bagging","Forest","Boosting"),
  rmses = c(lin.rmse.tst,tree.rmse.tst,bag.rmse.tst,
              forest.rmse.tst,boost.rmse.tst)))
```

# Conclusion
#### Linear Regression has the lowest RMSE, it performs best out of all. 